---
title: "gglot 1"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges) 
```

## create the weather data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
recode: rename the station code. 

## create a gglot

```{r}
ggplot(weather_df,aes(x = tmin, y =tmax)) + geom_point()

```

## alternate way of making this plot
start with a data, then do others, such as ggplot, filter..... 
```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) + geom_point()
```

saving initial plots, mostly don't do this
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY")
```
scatter plot
```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
scatterplot
```

adding color 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.4 )
```
alpha: degree of transparent (0-1). 
 
why do `aes` positions matter?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.4 ) + 
  geom_smooth()
 ##   geom_smooth(se = false) is used to turn off the line 
 ##  ggplot(aes(x = tmin, y = tmax, color = name)), if i put the color under gglot, both geom_point and geom_smooth will have the same color.(the color will be applied everywhere.)
```
the central line is used to describe the trend. 

facet!
separate into three figures!!!!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = 0.4 ) + 
  geom_smooth(se = FALSE) +
  facet_grid(~name)
  ## separate into three figures!!!!
``` 

temperature over the day of year in three locations, with precipitation. 
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.4) +
  geom_smooth(se = FALSE, size = 2)
```








