---
title: "ggplot 2"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges) 
library(patchwork) 
devtools::install_github("thomasp85/patchwork")
```

## create the weather data

```{r load_data, cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
recode: rename the station code. 
cache: save the outcome and don't have to be re downloaded again. 

## create a gglot
basic scatter plots
```{r}
ggplot(weather_df,aes(x = tmin, y =tmax)) + geom_point()

```

## alternate way of making this plot
start with a data, then do others, such as ggplot, filter..... 

Advanced Scatter plots
```{r}
weather_df %>% 
  ggplot(aes(x =tmin, y = tmax)) + geom_point()
```

saving initial plots, mostly don't do this
```{r}
weather_df %>% 
  filter(name == "CentralPark_NY")
```
scatter plot
```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
scatterplot
```

adding color 
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.4 )
```
alpha: degree of transparent (0-1). 
 
why do `aes` positions matter?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.4 ) + 
  geom_smooth()
 ##   geom_smooth(se = false) is used to turn off the line 
 ##  ggplot(aes(x = tmin, y = tmax, color = name)), if i put the color under gglot, both geom_point and geom_smooth will have the same color.(the color will be applied everywhere.)
```
the central line is used to describe the trend. 

facet!
separate into three figures!!!!
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = 0.4 ) + 
  geom_smooth(se = FALSE) +
  facet_grid(~name)
  ## separate into three figures!!!!
``` 

temperature over the day of year in three locations, with precipitation. 
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.4) +
  ## we can ignore this step if we wanna have only geom_smooth
  geom_smooth(se = FALSE, size = 2)
```

2d density 

```{r}
# install. packages("hexbin") for geom_hex 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_bin2d() +
  ## same things, but little squares. 
  facet_grid(~name)
```

## more kinds of plots!!
univariate plots 
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram() + 
  facet_grid(~name)
```

density plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, color = name, fill = name)) + 
  geom_density(alpha = 0.3)
```

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```


```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(fill = "yellow")
```

ridge plots!!!
```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  ggridges::geom_density_ridges()
```

## saving a plot
```{r}
ggp_ridge_temp = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges()

ggsave("ggplot_temp_ridge.pdf", ggp_ridge_temp)
```

there is difference between how r is creating a plot and how that get embeded inside here
Setting the width of figure and fit them into the reports!!!!!!  
```{r, fig.width = 12}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(aes(color = name), alpha = 0.4 ) + 
  geom_smooth(se = FALSE) 
```

## making new plots
start with the old plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 0.5)
```

add label:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (c)",
    y = "Maximum Temp (c)",
    caption = "data from noaa via rnoaa package"
  )
```

x axis tick marks etc. 

```{r} 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (c)",
    y = "Maximum Temp (c)",
    caption = "data from noaa via rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15,-5,20),
    labels = c("-15C", "-5 still too cold)", "20")
    ) + 
    scale_y_continuous(
       trans = "sqrt", 
       position = "right"
    )
  ## scale_y_sqrt() does the same thing. 
``` 
ggplot have some default things. we can do something on scales ourselves. 

```{r}

```

## colors 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (c)",
    y = "Maximum Temp (c)",
    caption = "data from noaa via rnoaa package"
  )+
  scale_color_hue(
  name = "Weather Station",
  h = c(1, 250)
  )
```
(100,300) is the color hue range.  

viridis: better than built-in. 
```{r}
ggp_base =
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point (alpha = 1) +
  labs(
    title = "Temperature plot",
    x = "Minimum Temp (c)",
    y = "Maximum Temp (c)",
    caption = "data from noaa via rnoaa package"
  )+
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_base
```

```{r}
ggp_base + 
  theme_bw() +
  theme(legend.position =  "bottom")
```

```{r}
ggp_base + 
  theme_minimal() +
  theme(legend.position =  "bottom")  
```
order matters. 

where is the warning coming from?
check the missing values!!!! 
```{r}
weather_df %>%   
filter(is.na(tmin))
```

general rules applied everywhere in the R Markdown.

## separate into different dataset and make them into  different geom 

```{r}
## dataset 1
central_park = 
weather_df %>% 
  filter(name == "CentralPark_NY")

## dataset 2
waikiki =
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp)) + 
  geom_line(data = central_park)
```

brief aside about colors

```{r}
waikiki  %>% 
  ggplot(aes(x=date, y = tmax)) +
  geom_point(alpha = 0.5, color = "blue")
```
when want to give the points color that is not a variable in the dataset, we need to put it after geom_point to specify the color. 

## multi-panel plots 
```{r}
ggp_scatter = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_density =
  weather_df %>% 
  ggplot(aes(x = tmin)) + 
  geom_density()
 
ggp_box =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_boxplot()

(ggp_scatter + ggp_density )/ggp_box
```










































